---
title: "HIP Sampling on Polygon Example"
author: "Aidan McDonald"
date: "October 17, 2018"
output: "html_document"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document demonstrates the process for using SDraw to generate and plot a Halton Iterative Partitioning (HIP) sample from a SpatialPolygons object.

```{r import, echo=TRUE}
library(SDraw)
```

## HIP Sampling

In this example, we will use HIP to draw a spatially balanced sample of 25 points from the state of Washington. Usually, it is recommended to use a fine Halton lattice with $J = (8,5)$, but for ease of visualization we will use a coarser lattice with $J = (3,2)$.

```{r sampling}
data("WA")
n <- 25
J <- c(3,2)
S <- sdraw(WA, n, "HIP", J = J)
```

## Plotting

For HIP sampling, SDraw has a special plotting function that overlays the Halton lattice on the polygon.
```{r plot}
hip.plot.lattice(WA, J = J, sample = S)
```

## References

*Robertson BL, McDonald T, Price CJ, Brown JA (2017) Halton iterative partitioning: spatially balanced sampling via partitioning. Environ Ecol Stat 25:305-323*

